---
title: "exercise-05-challenge-2"
author: "John Hinkle"
format: html
self-contained: true
editor: visual
---

## Exercise 05, Challenge 2

## Step 1: Load data

```{r}
library(tidyverse)
z <- read.csv("https://raw.githubusercontent.com/difiore/ada-datasets/main/zombies.csv")
```

## Step 2: Calculate the population mean and standard deviation for each quantitative random variable

```{r}
pop_sd <- function(x){
  sqrt(sum((x - mean(x))^2) / length(x))
}



print(z_summary <- 
        z |> 
        select(-id) |>
  summarise(across(where(is.numeric), list( 
    mean = ~mean(.), 
    sd = ~pop_sd(.)
    ), .names = "{.col}_{.fn}")))
```

## Step 3: Box plots of quantitative random variables by gender

```{r}
library(ggplot2)

# Height
height_boxplot <- ggplot(z, aes(x=gender, y=height)) +
  geom_boxplot(color = "#BF5700")
height_boxplot

# Weight
weight_boxplot <- ggplot(z, aes(x=gender, y=weight)) +
  geom_boxplot(color = "#BF5700")
weight_boxplot

# Zombies killed
kills_boxplot <- ggplot(z, aes(x=gender, y=zombies_killed)) +
  geom_boxplot(color = "#BF5700")
kills_boxplot

# Education
education_boxplot <- ggplot(z, aes(x=gender, y=years_of_education)) +
  geom_boxplot(color = "#BF5700")
education_boxplot

# Age
age_boxplot <- ggplot(z, aes(x=gender, y=age)) +
  geom_boxplot(color = "#BF5700")
age_boxplot

```

## Step 4: Scatter plots of height and weight vs. age

It appears that there is a positive relationship between height vs. age, and weight vs. age. As you get older, you get taller and heavier. It is worth noting that the age axis only goes from 10 years old to 30 years old.

```{r}

# Height
print(height_scatter <- ggplot(z, aes(x= age, y=height, color=gender)) +
  geom_point())

# Weight
print(weight_scatter <- ggplot(z, aes(x= age, y=weight, color=gender)) +
  geom_point())
```

## Step 5: Histograms of each random quantitative variable

'zombies_killed' and 'years_of_education' both have non-normal distributions.

```{r}
# Height
hist(z$height)
# Weight
hist(z$weight)

# Zombies killed
hist(z$zombies_killed)

# Years of education
hist(z$years_of_education)

# Age
hist(z$age)
```

## Step 6:

```{r}
z_sample <- slice_sample(z, n=50)

print(z_sample_summary <- 
        z_sample |> 
        select(-id) |>
  summarise(across(where(is.numeric), list( 
    mean = ~mean(.), 
    sd = ~sd(.)
    ), .names = "{.col}_{.fn}")))
```

## Step 7:

```{r}

```

## Step 8:

```{r}

```

## Step 9:

```{r}

```

## Step 10:

```{r}

```
